Superphy: Real-time prediction of \textit{E. coli} phenotype using graph-based genome sequence storage and retrieval with the \textit{spfy} module

### 2017-11-16 paper-webserver.tex
STOP: about to merge into master, no real pt on a branch atm
EDIT: read-through of intro

% 2. Big Problem: but tools are for individual analysis
The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications. Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} virulence gene annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}. In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to mimick the results of traditional tests. These methods allow new WGS results to be viewed in the context of historical tests.

Comprehensive platforms that combine individual programs into a cohesive whole also exist. These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC. Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests {REF}.

@@

% 5. how we solved it
In this study, we present an update to the SuperPhy platform, called Spfy. Spfy aims to solve the problem of real-time analysis and merging of results from different analysis methods. By merging results and creating links between genome data, we are able to identifying relationships between all genomes sequenced in the past, present, and future. This graph-based result storage allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible enough to accomadate new analysis methods as they are developed.

@@

By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, with added support for big-data analyses. Subtyping options are ..., pan-genome generation ..., group comparisons via Fisher's, ML, .... for E.coli. These tasks are divided into subtasks, and distributed across a built-in task queue. Subtyping results are then decomposed into graph structures and stored within a larger graph database. By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups. Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.

Future work will focus on adding additional analyses modules, using machine learning and artificial neural networks to aid genotype to phenotype predictions, and supporting different species. While the integrated approach or storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA.

The website and source code are available at \url{https://lfz.corefacility.ca/superphy/spfy/}. \par

@@

% para covering ectyper & RGI -- don't really need this
% RGI has its own paper and was in the previous SuperPhy paper
% ectyper will get its own paper

% para on phylotyper
Stx typing looks at ...
Phylotyper constructs phylogenetic trees and uses ancestral reconstruction to determine Stx-types ... others.
The software can also be used to generate trees and determine closely related strains.

>>>>

% 2. Big Problem: but tools are for individual analysis
The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications. Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} for virulence factor (VF) annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}. In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to replicate the results of traditional wet-lab tests; this allows new WGS results to be viewed in the context of historical tests.

Comprehensive platforms that combine individual programs into a cohesive whole also exist. These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC \cite{wattam2013patric}. Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests \cite{swaminathan2001pulsenet}.

@@

% 5. how we solved it
In this study, we present an update to the SuperPhy platform, called Spfy. Spfy aims to solve the problem of real-time analysis and merging of results from different analysis methods. By merging results and creating links between genome data, Spfy is able to identify relationships between all genomes sequenced in the past, present, and future. This graph-based result storage allows retrospective comparisons as more genomes are sequenced or populations change, and is flexible enough to accomadate new analysis methods as they are developed.

@@

By supporting multiple \textit{in-silico} subtyping options, the platform functions similar to a reference laboratory, with added support for big-data analyses. Subtyping options for E.coli are O-antigen, H-antigen, Shiga-toxin 1 (Stx1), Shiga-toxin 2 (Stx2), and Intimin typing, VF and AMR annotation, pan-genome generation, group comparisons via Fisher's, and machine-learning (ML) based predictions for OmniLog AMR assays. These tasks are divided into subtasks, and distributed across a built-in task queue. Results are also decomposed into graph structures and stored within a larger graph database. By integrating task distribution with graph storage, Spfy enables large-scale analyses, such as epidemiological associations between specific genotypes, biomarkers, host, source, and other metadata, and statistical significance testing of genome markers for user-defined groups. Currently, the platform has been tested with XXX genome files and result storage for XX analyses modules.

Future work will focus on adding additional analyses modules, using machine learning and artificial neural networks to aid genotype to phenotype predictions, and supporting different species. While the integrated approach or storing and retrieving results provides enormous benefits, the developed analyses modules are self-contained and can easily be integrated into existing platforms such as Galaxy \cite{goecks2010galaxy}, and IRIDA. The website and source code are available at \url{https://lfz.corefacility.ca/superphy/spfy/}.

@@

% para covering ectyper & RGI -- don't really need this
% RGI has its own paper and was in the previous SuperPhy paper
% ectyper will get its own paper
% phylotyper also has its own paper
